---
title: "IXIS_Challenge"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Add Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(magrittr)
library(DataExplorer)
library(skimr)
library(openxlsx)
library(viridis)
```

##Read in data from Github repository

```{r}
counts<-read_csv("https://raw.githubusercontent.com/sconnin/GA-Scenario/main/DataAnalyst_Ecom_data_sessionCounts.csv")
cart <- read_csv("https://raw.githubusercontent.com/sconnin/GA-Scenario/main/DataAnalyst_Ecom_data_addsToCart.csv")
```

## Conduct initial Exploration for Counts and Carts datasets


```{r}


introduce(counts) # basic data structure 
introduce(carts)

#assess number of unique entries in select columns

summarise(across(c(dim_browser, dim_deviceCategory, 
    dim_date),  n_distinct, rm.na=T)) 

# assess missing data  

plot_missing(counts) 
plot_missing(cart)

skim(counts) # Collect univariate stats  
skim(cart)

# assess distributions of discrete variables

plot_bar(counts)
plot_bar(cart)

# assess bivariate relationships (by device)

plot_bar(counts, with = "transactions")
plot_bar(counts, with = "sessions")
plot_bar(counts, with = "QTY")


```
## Clean & wrangle counts data and features as appropriate
```{r}


#1. Create Month, Day, Year columns for grouping, rename cols per instructions

sheet1<-counts%>%                 
    
    separate(dim_date,  c("Month", "Day", "Year"))%>%
    
    mutate(Month=recode(Month, '1' = 'January', '2' = 'February', '3' = 'March', 
        '4' = 'April', '5' = 'May', '6' = 'June', '7' = 'July', '8' = 'August', 
        '9' = 'September', '10' = 'October', '11' = 'November', 
        '12' = 'December'))%>%
    
    mutate(Year=recode(Year, '12' = '2012', '13' = '2013'))%>%
    
    rename(Transactions = transactions, Sessions = sessions, 
           Device = dim_deviceCategory, Browser = dim_browser)
    
#2. create ECR ratio column per instructions
    
sheet1%<>%mutate(ECR = round(Transactions/Sessions, 2))

#3. replace 0 with NA for ECR where Sessions = 0 to avoid inf in summary stats
    
sheet1%<>%mutate(ECR=ifelse(Sessions == 0, NA, ECR))
    
#3 remove non-value GA returns for Browser
    
sheet1%<>%filter(Browser != "(not set)" & Browser != "error" )
    
#4. order columns starting with Year and Month to enhance review of chronology
    
sheet1%<>%relocate(Year, .before = Browser)%>%
    relocate(Month, .after = Year)%>%
    relocate(Day, .after = Month)
    
#5. Aggregate data per instructions 

sheet1%<>%group_by(Month, Device)


```

## Clean & wrangle counts data and features as appropriate

```{r}

# Create Month, Day, Year columns for grouping

crt<-cart%>%
    
    rename(Year = dim_year, Month = dim_month, AddsToCart = addsToCart)%>%
    
    mutate(Month=recode(Month, '1' = 'January', '2' = 'February', '3' = 'March', 
                        '4' = 'April', '5' = 'May', '6' = 'June', '7' = 'July', '8' = 'August', 
                        '9' = 'September', '10' = 'October', '11' = 'November', 
                        '12' = 'December'))
    
#2. select last two months
    
crt%<>%tail(2)

```

# Combine and save datasets into xlsx worksheets

```{r}

#1. subset datasets in advance of joining to create second worksheet

AddsToCart<-crt$AddsToCart

sheet1b<-sheet1%>%
    filter(Year==2013)%>%
    filter(Month == 'May' | Month == 'June' )%>%
    group_by(Year, Month)%>%
    summarise_if(is.numeric, sum, na.rm = TRUE)

#2. combine datasets to create sheet2

sheet2<-cbind(AddsToCart, sheet1b)%>%
    rename(AddsToCart = ...1)%>%
    relocate(AddsToCart, .after = ECR)%>%
    pivot_wider(names_from = Month, values_from = c(Sessions, 
                                                    Transactions, QTY, ECR, AddsToCart))%>%
    
    mutate(Sess_Adiff = Sessions_June-Sessions_May, 
           Sess_RDiff = (Sessions_June-Sessions_May)/Sessions_June)%>%
    relocate(c(Sessions_June,Sess_Adiff, Sess_RDiff), .after = Sessions_May )%>%
    
    mutate(Trans_Adiff = Transactions_June-Transactions_May, 
           Trans_RDiff = (Transactions_June-Transactions_May)/Transactions_June)%>%
    relocate(c(Transactions_June,Trans_Adiff, Trans_RDiff), .after = Transactions_May)%>%
    
    mutate(QTY_Adiff = QTY_June-QTY_May, 
           QTY_RDiff = (QTY_June-QTY_May)/QTY_June)%>%
    relocate(c(QTY_June, QTY_Adiff, QTY_RDiff), .after = QTY_May )%>%
    
    mutate(ECR_Adiff = ECR_June-ECR_May, 
           ECR_RDiff = (ECR_June-ECR_May)/ECR_June)%>%
    relocate(c(ECR_June, ECR_Adiff, ECR_RDiff), .after = ECR_May )%>%
    
    mutate(Cart_Adiff = AddsToCart_June-AddsToCart_May, 
           Cart_RDiff = (AddsToCart_June-AddsToCart_May)/AddsToCart_June)%>%
    relocate(c(AddsToCart_June, Cart_Adiff, Cart_RDiff), 
             .after = AddsToCart_May)%>%
    
    rename(Sess_May=Sessions_May, Sess_June=Sessions_June, 
           Trans_May = Transactions_May, Trans_June = Transactions_June, 
           Cart_May = AddsToCart_May, Cart_June = AddsToCart_June)

# 3. write sheets to single xlxs file in home directory

excel <- list("Counts" = sheet1, "Two_Month_Compare" = sheet2)

write.xlsx(excel, file = "GA_challenge.xlsx")
```

## Compile basic univariate stats for worksheet1 with count data

```{r}

#Derive stats for the 12 month period

summary(sheet1)

#Derive stats grouping by month, device, and browser

(cnt_month<-sheet1%>%
    group_by(Month, Device, Browser)%>%
    summarise(
        tot_sess = sum(Sessions),
        mn_sess = mean(Sessions, na.rm=T), 
        sd_sess = sd(Sessions, na.rm=T),
        tot_trans = sum(Transactions),
        mn_trans=mean(Transactions, na.rm=T), 
        sd_trans = sd(Transactions, na.rm=T),
        mn_ecr = round(mean(ECR, na.rm=T),3), 
        sd_ecr=round(sd(ECR, na.rm=T), 3)))



```
## Create data graphics for analysis and presentation

```{r}
#1. Plot total Sessions by Month

s<-cnt_month%>%
    ggplot(aes(x=Month, y=tot_sess, fill=Device))+
    geom_col(alpha =  0.8)+
    scale_x_discrete(limits = rev(levels(levels)))+
    scale_fill_viridis_d(option="E")+
    ggtitle("Sessions Per Month") + 
    theme(plot.title = element_text(
        size = 10,
        lineheight = .9,
        family = "Times",
        colour = "black"))+
    ylab("Total Count")+
    theme(axis.text.y = element_text(
        size=8))+
    xlab("Month")+
    theme(axis.text.x = element_text(
        size=8))+
    theme_classic()

(sess<-s+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)))

#2. Plot total transactions by month

t<-cnt_month%>%
    ggplot(aes(x=Month, y=tot_trans, fill=Device))+
    geom_col(alpha =  0.8)+
    scale_x_discrete(limits = rev(levels(levels)))+
    scale_fill_viridis_d(option="E")+
    ggtitle("Transactions Per Month") + 
    theme(plot.title = element_text(
        size = 10,
        lineheight = .9,
        family = "Times",
        colour = "black"))+
    ylab("Total Count")+
    theme(axis.text.y = element_text(
        size=8))+
    xlab("Month")+
    theme(axis.text.x = element_text(
        size=8))+
    theme_classic()

(trans<-t+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)))

#3. Plot Average eConversion per Month

c<-cnt_month%>%
    ggplot(aes(x=Month, y=mn_ecr, fill=Device))+
    geom_col(alpha =  0.8)+
    scale_x_discrete(limits = rev(levels(levels)))+
    scale_fill_viridis_d(option="E")+
    ggtitle("Average Conversion Rate Per Month") + 
    theme(plot.title = element_text(
        size = 10,
        lineheight = .9,
        family = "Times",
        colour = "black"))+
    ylab("Average")+
    theme(axis.text.y = element_text(
        size=8))+
    xlab("Month")+
    theme(axis.text.x = element_text(
        size=8))+
    theme_classic()

(conv<-c+theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)))

#4. Plot device & Browser in relation to Transactions

dev_trans<-counts%>%
    select(dim_browser, dim_deviceCategory, transactions)%>%
    rename(Browser = dim_browser, Device = dim_deviceCategory,
           Transactions=transactions)%>%
    group_by(Device)%>%
    summarise(across(where(is.numeric), 
                     list(mean = mean, sum = sum)))

(dev_trans%>%
    mutate_at(vars(contains('Device')),as.factor)%>%
    ggplot(aes(x=reorder(Device, Transactions_sum), 
               y =Transactions_sum))+
    geom_col(aes(fill=Device), width=0.4, show.legend = FALSE)+
    coord_flip()+
    scale_fill_viridis_d(option="E")+
    ggtitle("Total Transactions by Device") + 
    theme(plot.title = element_text(
        size = 10,
        lineheight = .9,
        family = "Times",
        colour = "black",
        vjust = 5))+
    ylab("Total")+
    theme(axis.text.y = element_text(
        size=8))+
    xlab("Device")+
    theme(axis.text.x = element_text(
        size=8))+
    theme_classic())

#5. Plot browser vs device

(brows_desk<-counts%>%
    filter(dim_deviceCategory == 'desktop')%>%
    select(dim_browser,  transactions)%>%
    rename(Browser = dim_browser, Transactions=transactions)%>%
    group_by(Browser)%>%
    summarise(across(where(is.numeric), list(sum = sum)))%>%
    filter(Transactions_sum >1000)%>%
    arrange(desc(Transactions_sum))%>%
    mutate_at(vars(contains('Browser')),as.factor)%>%
    
    ggplot(aes(x=reorder(Browser, Transactions_sum), 
               y =Transactions_sum))+
    geom_col(aes(fill=Browser), width=0.4, show.legend = FALSE)+
    coord_flip()+
    scale_fill_viridis_d(option="E")+
    ggtitle("Transactions by Desktop Browser:\n Where Total > 1000") + 
    theme(plot.title = element_text(
        size = 10,
        lineheight = .9,
        family = "Times",
        colour = "black",
        vjust = 5))+
    ylab("Total")+
    theme(axis.text.y = element_text(
        size=8))+
    xlab("Browser")+
    theme(axis.text.x = element_text(
        size=8))+
    theme_classic())

#6. Plot transactions vs sessions by device

(trans_sess<-sheet1%>%
    filter(!(ECR > 1))%>%
    ggplot(aes(x=Sessions, y=Transactions))+
    geom_point(aes(color=Device))+
    ggtitle("Transactions by Session Count and Device")+ 
    theme(plot.title = element_text(
        size = 10,
        lineheight = .9,
        family = "Times",
        colour = "black",
        vjust = 5))+
    ylab("Total Transactions")+
    theme(axis.text.y = element_text(
        size=6))+
    xlab("Total Sessions")+
    theme(axis.text.x = element_text(
        size=6))+
    theme_classic())

#7. Plot items added to cart by month - refer back to raw data

# wrangle and factor data for graph creation


Adds<-cart%>%
    rename(Year = dim_year, Month = dim_month, AddsToCart = addsToCart)%>%
    
    mutate(Month=recode(Month, '1' = 'January', '2' = 'February', '3' = 'March','4' = 'April', '5' = 'May', 
        '6' = 'June', '7' = 'July', '8' = 'August','9' = 'September', '10' = 'October', '11' = 'November',
        '12' = 'December'))


Adds$Month<-as.factor(Adds$Month)

Adds$Month<- ordered(Adds$Month, levels = c('December', 'November', 'October','September', 'August',  'July','June','May','April', 'March','February', 'January'))

# plot graph

(Adds%>%
    ggplot(aes(Month, AddsToCart))+
    geom_col(fill="light blue", alpha=.9)+
    ggtitle("Items Added to Cart Each Month")+ 
    theme(plot.title = element_text(
        size = 8,
        lineheight = .9,
        family = "Times",
        colour = "black",
        vjust = 5))+
    ylab("Total")+
    theme(axis.text.y = element_text(
        size=6))+
    xlab("Month")+
    theme(axis.text.x = element_text(size=6))+
    coord_flip()+
    theme_classic())
```

